Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow

### 2012-07-27 Text/introduction.tex
Went to editor.

\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{bm}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{tabularx}
\usepackage{leftidx}

@@

\title{Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow}

\begin{document}
\maketitle
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in a slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circular circumferences as
is often done in the literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the scalar concentration is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration. %, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where the tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration for the range of capillary numbers $0.1<Ca<1.0$.
We show that the time-dependent average concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one unified equation. %,
% Same question applies.
%Eq. \ref{main:main:main}. 
Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions of \citet{vanbaten-circular}. %The work is of use to people performing mass transfer numerical simulations for bubble train flow in microchannels within the lattice Boltzmann context.
\end{abstract}

\section{Introduction}
\label{intro}
Monolith reactors have recently been getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}.  These reactors usually operate in
 the Bretherton-Taylor regime \cite{bretherton,taylor} which describes a flow
 of equally sized, long air bubbles through a liquid medium, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by the dominance of surface tension over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to the dominance of surface tension, the flow exhibits advantageous
 properties which cannot be achieved
in its macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls strongly enhance mass
transfer from gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}.
 Moreover, the low slip velocity between gas and liquid is utilized in experiments to measure liquid velocity \cite{taylor}:  bubbles
 travelling with approximately the same velocity as liquid can be captured with a camera. These properties explain why nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, chemical reactors for hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. The extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

This work is focused on  gas to liquid mass transfer  for the two-dimensional Bretherton/Taylor flow. A good understanding of mass transfer and how it depends on 
parameters such as the capillary number, the Reynolds number, and slug and bubble lengths allows to properly
manufacture a microchannel with properties necessary to ensure that chemical
reactions are performed in the best possible manner. The mass transfer coefficient is defined as the flux from the 
surface divided by the difference of the imposed concentration and the characteristic concentration in the domain.
 The  concentration distribution in the domain is prescribed by underlying hydrodynamics fields. 
 For example, experimental studies \cite{yue-mass,bercic-mass} show a complex dependency of the mass transfer coefficient on flow parameters:
 bubble and slug lengths, and bubble velocity, which  in turn relate to the capillary number $Ca$ and the Reynolds number $Re$. 
\citet{yue-mass} established an experimental correlation for the volumetric mass transfer coefficient for a bubble train as
a function of the diffusion coefficient, slug and bubble lengths, and bubble velocity: 
\begin{equation}
\vol =\frac{2}{d_h} \Bigl(\frac{D
\ububble}{\lbubble+\lslug}\Bigr)^{0.5}
\Bigl(\frac{\lbubble}{\lbubble+\lslug}\Bigr)^{0.3},
\end{equation}
where $\vol$ is the volumetric mass transfer coefficient, $d_h$ is the hydraulic
diameter, $\lbubble$ is the bubble length, $\lslug$ is the slug distance (between bubbles),
$\ububble$ is the bubble velocity, and $D$ is the diffusion coefficient. 
%This correlation is within $10\%$ accuracy for experimental
%results for the microchannels with rectangular/square crosssections for the following range of
%parameters: bubble velocity $0.4\,\mathrm{m/s}<\ububble<2\,\mathrm{m/s}$, bubble length
%$1.4<\frac{\lbubble}{d_h}<6.3$, slug length $1<\frac{\lslug}{d_h}<3.2$,
%hydrodynamic diameter $d_h = 400 \mathrm{\mu m}$.

@@

% TODO: move this later so that there is no forward reference to eq.4? It's ok to have it here
Since the mass transfer coefficient is defined in terms of a mass flux through a certain area, Eq. \ref{eq:main:definition},
analytical estimates \cite{kreutzer-overview,irandoust} are based on a decomposition of the bubble shape. The mass
 transfer coefficient is calculated through two separate contributions from two planar films and two hemicircles. For both 
contributions the Higbie penetration theory \cite{higbie} is utilized, which states that the mass transfer coefficient
 for a simple flow geometry depends on the average time a liquid packet interacts with a geometrical feature. It 
can be calculated as $\sqrt{\frac{\pi D}{t_{\mathrm{char}}}}$, where $t_{\mathrm{char}}$ is the interaction 
time. As an example of the application of the Higbie penetration theory,
 the mass  transfer coefficient 
for the flow of bubbles between parallel plates  is calculated as (similarly to the work of \citet{vanbaten-circular}):
\beq
k_L=2 \sqrt{\frac{\pi D}{t_{\mathrm{film}}}}+2 \sqrt{\frac{\pi D}{t_{\mathrm{circle}}}},
\feq
where $t_{\mathrm{film}}=\frac{\lfilm}{\ububble}$ stands for the interaction time of liquid traveling below the planar part
 of the bubble, $t_{\mathrm{circle}}=\frac{\pi R_{\mathrm{circle}}}{\ububble}$ is the time during which the liquid in the slug travels the distance of half the bubble cap circumference.

Despite their simplicity, such analytical expressions work well for flows with low
capillary numbers $Ca<0.1$ \cite{bercic-mass} where the bubble shape is
symmetrical and can be approximated with good precision. Moreover, because of the
hydrodynamic pattern in the slug (i.e. presence of a vortex in slug), one can
estimate the time for a fluid batch to travel the whole circumference.   However,
with the increase of the capillary number the situation changes significantly -- the
symmetrical bubble shape is lost and the bubble shape resembles a bullet \cite{kuzmin-binary2d}.
For flows with $Ca>0.7$ there is also no vortex in the liquid slug. In this case the Higbie theory fails
to estimate the contribution from bubble caps, which explains the need to turn to numerical simulations where all
hydrodynamic fields as well as complex bubble shapes are taken into account.

Typical numerical studies of mass transfer
\cite{kreutzer-overview,vanbaten-circular} do not consider the simulation of
bubble shapes for $Ca>0.1$. The usual simulation of mass transfer is performed as
follows: 
\begin{description}
 \item[I] The bubble shape is calculated either
through analytical correlations \cite{bretherton} or experimental correlations
\cite{cerro-bubble-train} without directly resolving bubble shapes through
multiphase simulations.  The expressions for bubble shapes are available only
for flows with capillary number $Ca<0.1$.
  \item[II] Hydrodynamic fields
are then obtained by performing simulations of one-component flow around the
bubble by imposing the bubble velocity on the walls. Thus, the simulations are
performed in the reference frame moving with the bubble. A stress-free
condition is imposed at the bubble surface.
  \item[III]  The mass transfer
simulations are performed in the reference frame moving with the bubble. The
saturation concentration is imposed at the bubble surface. Only one unit cell
containing a single bubble is used for simulations. Periodic concentration boundary conditions are
utilized to determine the volumetric mass transfer coefficient, which is
calculated through the following equation \cite{vanbaten-circular}:
\begin{equation} \label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\cinoutlet\rangle}
\frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\ cell\ volume}},
\end{equation}
 where $\langle\cinoutlet(t)\rangle=\int{C \uinoutlet
\mathrm{d}A}/\int{\uinoutlet\mathrm{d}A}$ is the space-averaged inlet/outlet (periodic boundary conditions)
concentration as a function of time. Therefore, in terms of the mass transfer definition, $\langle\cinoutlet(t)\rangle$ plays the role of the 
characteristic concentration.
%This is the characteristic concentration in the definition, Eq. \ref{eq:main:definition}.  
The time-averaged
concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference
between the overall average concentration in the whole domain
($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$) at time $t_1$ and at time
$t_2$ divided by the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and experimental correlations of \citet{bercic-mass}
% TODO: was good where?
was good. 
 \end{description}

The presented numerical approaches \cite{vanbaten-circular,kreutzer-overview}
can be criticized on a number of points. They mainly
originate from the bubble shape approximation, which is taken to be symmetrical, i.e.
consisting of two hemispheres and film for the case of flow in circular
capillaries. This is valid  for small capillary numbers only ($Ca<0.1$). As previously
discussed, for such capillary numbers the tracer is well mixed in the slug and 
the choice of the characteristic concentration needed for the mass transfer
coefficient, Eq. \ref{eq:main:definition}, is obvious. With minimal
differences in the results, it can either be the averaged concentration in the liquid slug
or the inlet/outlet space-averaged concentration. The latter is used in the formulation
of \citet{vanbaten-circular} presented above.

While it is clear that periodic boundary conditions can be employed for the calculation
 of hydrodynamic fields, the same does not apply to the mass transfer coefficient
simulations. Experimental correlations \cite{bercic-mass} show that the
concentration along the streamwise direction changes exponentially with distance.
Mass transfer simulations however, are made only for one unit cell
using periodic boundary conditions with the same concentration at the inlet
and at the outlet. The question is how single unit cell simulation corresponds to
experimental measurements arises where concentration difference is measured at
the distances of at least a few unit cells \cite{bercic-mass}. In
other words, one needs to understand how the discrete one unit cell simulation
corresponds to the continuous picture in experiments where one does not
distinguish discrete bubbles but takes measurements of concentration at different
locations.

Addressing situations for a rich number of hydrodynamic patterns, shapes, and
effects of bubble lengths, etc for bubble train flows, we feel that there is a need to examine 
carefully the strategies and assumptions behind the numerical calculations
of the mass transfer coefficient. We aim at establishing clear
procedures as to how properly obtain the mass transfer coefficient via a study
of different boundary conditions and different definitions of the characteristic concentration.
The case we want to examine is a two-dimensional bubble train flow between parallel plates.
We address the following issues:
 \begin{description}
 \item[I] Applicability of periodic boundary
conditions to determine the mass transfer coefficient when the vortex in the slug
disappears, i.e. $Ca>0.7$.
 \item[II] Validity of the inlet/outlet-averaged or domain-averaged concentrations as
characteristic concentrations in the definition of the mass transfer
coefficient.
 \item[III] Translation of the continuous experimental picture to
numerical simulations of a few unit cells, the issue of correspondence
between space averages (simulations \cite{vanbaten-circular}) and time averages (experiment).
% \item[IV] Influence of the hydrodynamic pattern in flows with $Ca>0.7$ on the mass transfer.
\end{description}
In addition, at the end of the manuscript we present results of the dependence of the volumetric
mass transfer coefficient on the Peclet number that we compare with analytical \cite{irandoust} and experimental correlations
\cite{yue-mass}.  The thorough determination of the mass transfer coefficient and associated Sherwood number as a
function of other non-dimensional parameters such as gas holdup, bubble/slug
lenghts, and the capillary number is left for future studies.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_hydro.eps}
\caption{Simplified sketch of the bubble motion. Using periodic
conditions for the velocity field is natural, but needs evaluation for mass
transfer.
\label{fig:benchmark:hydro}}
\end{figure}

@@

The mass transfer problems in the lattice Boltzmann framework were mainly
addressed in a series of works of Ginzburg and co-authors
\cite{ginzburg-main,ginzburg-boundary-conditions,ginzburg-saturated-flow}.
In contrast to these works whose focus was on simulating the
advection-diffusion equation via the lattice Boltzmann framework, we concentrate
on the application side.  One should also mention the work of \citet{inamuro-scalar-boundary}
about heat and mass transfers in porous media and the work of
\citet{jos-mass} simulating lateral mixing in cross-channel flow.  The last two
works are focused on problems of homogeneous nature and do not provide
guidance as to how to obtain the mass transfer coefficient for heterogeneous cases.

The paper is organized as follows. We start with definitions of the volumetric mass transfer coefficient and apply them 
for the bubble train flow to derive expressions to connect the space- and time-averages.  Then, the lattice
Boltzmann model used to simulate mass transfer is presented and followed by benchmarks. Finally,
numerical simulations of various boundary conditions and simulations spanning a few unit cells for
different hydrodynamic patterns are presented to establish the procedure to determine the
volumetric mass transfer coefficient. The comparison with analytical correlations is also presented.

\section{Mass transfer definitions}
By definition, the mass transfer coefficient from a surface with an imposed constant
concentration $\cbubble$ is:
\beqal
\label{eq:main:definition}
&k_L=\frac{\dot{m}}{P \Delta C},\ \Delta C= \cbubble - \cmedium,
\feqal
where $\dot{m}$ is the mass exchange rate $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, and $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of 
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at an infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain or the flux-averaged
concentration at the inlet or outlet, etc. Thus, one needs to establish a clear definition of  $\Delta C$ to determine the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

@@

%Another issue is that if the bubble is long enough then the film saturates with the tracer
%concentration and its influence on mass transfer can be negligible \cite{vanbaten-circular}.
Overall, mixing patterns of the film and liquid slugs are of great importance for the analytical
estimation of mass transfer \cite{yue-mass}. However, the assumptions usually taken for 
mass transfer calculations are small capillary numbers and certain mixing patterns such as to help to
estimate the mass transfer using the penetration theory of \citet{higbie}.

In comparison with analytical calculations and simplifications, the numerical approach can take into
account the complex mixing patterns and geometries. However, there are challenges as  how to
mimic the continuous picture where reacting medium is moving with bulk velocity $U=\ugas+\uliq$  as it is done in
experiments. Thus, the questions indicated in Section \ref{intro} arise.  The next section gives more
 details about numerical simulations.

\subsection{Numerical simulations}
\label{section:cases}
Ideally one wants to mimic the continuous picture as it is seen in experiments.
Thus, mass transfer simulations for a number of unit cells are needed. As was
indicated above, there are two approaches towards it -- either to simulate the
bubble train and then to measure concentration along the pipe, Eq.~\ref{main:mass:transfer:expression},
or to transfer to the reference frame moving with the bulk
velocity $U$ and conduct the same measurements. However, both methods require tracking of
moving bubbles which is complicated from the numerical point of view. Therefore, one needs to come
up with a simple and smaller domain for calculations of the mass transfer coefficient, which closely mimics the
continuous picture by an infinite number of  separated bubbles.

@@

\section{Validation}
As was discussed earlier, analytical correlations for the mass transfer coefficient
are derived  as two separate contributions: the mass transfer from two
half circles and the mass transfer from the film. We will examine these
mass transfer cases closely with the help of the lattice Boltzmann method and compare them
against analytical
solutions. The next sections will give a short introduction to the lattice Boltzmann method and present benchmark results.

@@

velocities $2,4,6,8,10,15,20,40$ times larger than original velocities. The velocities and

@@

\ref{table:scaling:peclet} (top to bottom: 6 scalings, 5 scalings, 2 scalings, 1 scaling, 1 scaling). \label{fig:contours:scaling:peclet}}

@@

\section{The Poiseuille parabolic profile}

>>>>

\documentclass[preprint,12pt]{elsarticle}
\usepackage{lineno}
\journal{Chemical Engineering Journal}

\usepackage{amsmath,amssymb}
\usepackage{bm}
%\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{color}
\usepackage{tabularx}
\usepackage{leftidx}

@@

\begin{document}
\begin{frontmatter}
\title{Lattice Boltzmann study of mass transfer for two-dimensional Bretherton/Taylor bubble train flow}
\author[uofa]{A.~Kuzmin\corref{cor1}}
\cortext[cor1]{Corresponding author}
\ead{kuzmin@ualberta.ca}
\author[us]{M.~Januszewski}
\ead{michalj@gmail.com}
\author[schlum]{D.~Eskin}
\ead{deskin@slb.com}
\author[schlum]{F.~Mostowfi}
\ead{fmostowfi@slb.com}
\author[uofa]{J.J.~Derksen}
\ead{jos@ualberta.ca}
\address[uofa]{Chemical and Materials Engineering, University of Alberta\\ 7th Floor, ECERF, 9107
116 St, Edmonton, Alberta, T6G
2V4 Canada}
\address[us]{Insitute of Physics, University of Silesia, 40-007 Katowice, Poland}
\address[schlum]{Schlumberger DBR Technology Center\\ 9450 17 Ave NW, Edmonton, Alberta, T6N 1M9
Canada}
\begin{abstract}
This work presents a procedure for the determination of the volumetric mass transfer
coefficient in the context of lattice Boltzmann simulations for the Bretherton/Taylor bubble train
flow for capillary numbers $0.1 < Ca < 1.0$. We address the case where the hydrodynamic pattern changes from having a
vortex in a slug ($Ca<0.7$) to not having it ($Ca>0.7$) \cite{giavedoni-numerical}.  In the latter case
the bubble shape is asymmetric and cannot be approximated through flat surfaces and circular circumferences as
is often done in the literature \cite{vanbaten-circular,kreutzer-overview}.  When the vortex is present
in the slug, the scalar concentration is well mixed and it is common to use periodic boundary conditions and the inlet/outlet-averaged concentration
% Do we really want equation references in the abstract?
as the characteristic concentration. %, Eq. \ref{eq:main:definition}.
The latter is not valid for flows where the tracer is not well mixed, i.e. $Ca>0.7$.  We therefore
examine various boundary conditions (periodic, open, open with more than 1 unit cell) and definitions of the characteristic concentration to estimate mass transfer coefficients  for the range of capillary numbers $0.1<Ca<1.0$.
We show that the time-dependent volume averaged concentration taken as the characteristic concentration produces the most robust results and
that all strategies presented in the literature are extreme limits of one unified equation. %,
% Same question applies.
%Eq. \ref{main:main:main}. 
Finally, we show good agreement of simulation results for different Peclet numbers
with analytical predictions of \citet{vanbaten-circular}. %The work is of use to people performing mass transfer numerical simulations for bubble train flow in microchannels within the lattice Boltzmann context.
\end{abstract}
\begin{keyword}
Mass Transfer \sep Taylor/Bretherton bubble train flow \sep Multiphase flow \sep Lattice Boltzmann
method
\sep Binary liquid model \sep Flow in microchannels with parallel plates 
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}
\linenumbers

\section{Introduction}
\label{intro}
Monolith reactors have recently been getting more attention as a promising alternative to slurry
reactors and trickle bed reactors \cite{kreutzer-overview,bercic-mass}.  These reactors usually operate in
 the Bretherton-Taylor regime \cite{bretherton,taylor} which is a flow
 of equally sized, long air bubbles through a liquid medium, see
Fig. \ref{fig:benchmark:hydro}. This flow regime
is characterized by the dominance of surface tension over inertia and viscous effects, and by
comparatively small gas flow velocities \cite{yue-mass}. Due to the dominance of surface tension, the flow exhibits advantageous
 properties which cannot be achieved
in its macroscopic counterparts: liquid thin films \cite{bretherton} between bubbles and walls strongly enhance mass
transfer from gas and walls to liquid; the plug flow regime occurring in monolith reactors allows to perform chemical reactions in slugs only \cite{kreutzer-overview}.
 Moreover, the low slip velocity between gas and liquid is utilized in experiments to measure liquid velocity \cite{taylor}:  bubbles
 travelling with approximately the same velocity as liquid can be captured with a camera. These properties explain why nowadays one can find a large number of applications
of the Bretherton-Taylor bubble train flow: continuous flow analyzers to measure liquid velocity, chemical reactors for hydrogenation of nitroaromatics, 2-ethyl-hexenal, Fischer-Tropsch synthesis, etc. The extensive reviews of \citet{kreutzer-overview,
gupta-review,yue-mass} cover a significant number of applications.

This work is focused on  gas to liquid mass transfer  for the two-dimensional Bretherton/Taylor flow. A good understanding of mass transfer and how it depends on 
parameters such as the capillary number, the Reynolds number, and slug and bubble lengths allows to properly
manufacture a microchannel with properties necessary to ensure that chemical
reactions are performed in the best possible manner. The mass transfer coefficient is defined as the flux from the 
gas-liquid interface divided by the difference of the imposed concentration and the characteristic concentration in the domain.
 The  concentration distribution in the domain is prescribed by underlying hydrodynamics fields. 
 For example, experimental studies \cite{yue-mass,bercic-mass} show a complex dependency of the mass transfer coefficient on flow parameters:
 bubble and slug lengths, and bubble velocity, which  in turn relate to the capillary number $Ca$ and the Reynolds number $Re$. 
\citet{yue-mass} established an experimental correlation for the volumetric mass transfer coefficient for a bubble train as
a function of the diffusion coefficient, slug and bubble lengths, and bubble velocity: 
\begin{equation}
\vol =\frac{2}{d_h} \Bigl(\frac{D
\ububble}{\lbubble+\lslug}\Bigr)^{0.5}
\Bigl(\frac{\lbubble}{\lbubble+\lslug}\Bigr)^{0.3},
\end{equation}
where $\vol$ is the volumetric mass transfer coefficient, $d_h$ is the hydraulic
diameter, $\lbubble$ is the bubble length, $\lslug$ is the slug distance (between bubbles),
$\ububble$ is the bubble velocity, and $D$ is the diffusion coefficient. 
%This correlation is within $10\%$ accuracy for experimental
%results for the microchannels with rectangular/square crosssections for the following range of
%parameters: bubble velocity $0.4\,\mathrm{m/s}<\ububble<2\,\mathrm{m/s}$, bubble length
%$1.4<\frac{\lbubble}{d_h}<6.3$, slug length $1<\frac{\lslug}{d_h}<3.2$,
%hydrodynamic diameter $d_h = 400 \mathrm{\mu m}$.

@@

% TODO: move this later so that there is no forward reference to eq.4? It's ok to have it here
Since the mass transfer coefficient is defined in terms of a mass flux through a certain area, Eq. \ref{eq:main:definition},
analytical estimates \cite{kreutzer-overview,irandoust} are based on a decomposition of the bubble surface in parts. The mass
 transfer coefficient is calculated through two separate contributions from two planar films and two hemicircles. For both 
contributions the Higbie penetration theory \cite{higbie} is utilized, which states that the mass transfer coefficient
 for a simple flow geometry depends on the average time a liquid packet interacts with a geometrical feature. It 
can be calculated as $\sqrt{\frac{\pi D}{t_{\mathrm{char}}}}$, where $t_{\mathrm{char}}$ is the interaction 
time. As an example of the application of the Higbie penetration theory,
 the mass  transfer coefficient 
for the flow of bubbles between parallel plates  is calculated as (similar to the work of \citet{vanbaten-circular}):
\beq
k_L=2 \sqrt{\frac{\pi D}{t_{\mathrm{film}}}}+2 \sqrt{\frac{\pi D}{t_{\mathrm{circle}}}},
\feq
where $t_{\mathrm{film}}=\frac{\lfilm}{\ububble}$ stands for the interaction time of liquid traveling next to the planar part
 of the bubble, and $t_{\mathrm{circle}}=\frac{\pi R_{\mathrm{circle}}}{\ububble}$ is the time during which the liquid in the slug travels the distance of half the bubble cap circumference.

Despite their simplicity, such analytical expressions work well for flows with low
capillary numbers $Ca<0.1$ \cite{bercic-mass} where the bubble shape is
symmetrical and can be approximated with good precision. Moreover, because of the
hydrodynamic pattern in the slug (i.e. presence of a vortex in slug), one can
estimate the time for a fluid batch to travel the whole circumference.   However,
with the increase of the capillary number the situation changes significantly -- the
symmetrical bubble shape is lost and the bubble resembles a bullet \cite{kuzmin-binary2d}.
For flows with $Ca>0.7$ there is also no vortex in the liquid slug. In this case the Higbie theory fails
to estimate the contribution from bubble caps, which explains the need to turn to numerical simulations where all
hydrodynamic fields as well as complex bubble shapes are taken into account.

Typical numerical studies of mass transfer
\cite{kreutzer-overview,vanbaten-circular} do not consider the simulation of
bubble shapes for $Ca>0.1$. The usual simulation of mass transfer is performed as
follows: 
\begin{description}
 \item[I] The bubble shape is calculated either
through analytical correlations \cite{bretherton} or experimental correlations
\cite{cerro-bubble-train} without directly resolving bubble shapes through
multiphase simulations.  The expressions for bubble shapes are available only
for flows with capillary number $Ca<0.1$.
  \item[II] Hydrodynamic fields
are then obtained by performing simulations of one-component flow around the
bubble by imposing the bubble velocity on the channel walls. Thus, the simulations are
performed in the reference frame moving with the bubble. A stress-free
condition is imposed at the bubble surface.
  \item[III]  The mass transfer
simulations are performed in the reference frame moving with the bubble. The
saturation concentration is imposed at the bubble surface. Only one unit cell
containing a single bubble is used for simulations. Periodic concentration boundary conditions are
utilized to determine the volumetric mass transfer coefficient, which is
calculated through the following equation \cite{vanbaten-circular}:
\begin{equation} \label{main:simulation:equation}
\vol=\frac{\mathrm{\overline{Flux}}}{\cbubble-\langle\cinoutlet\rangle}
\frac{\mathrm{bubble\ surface\ area}}{\mathrm{unit\ cell\ volume}},
\end{equation}
 where $\langle\cinoutlet(t)\rangle=\int{C \uinoutlet
\mathrm{d}A}/\int{\uinoutlet\mathrm{d}A}$ is the space-averaged inlet/outlet (periodic boundary conditions)
concentration as a function of time. Therefore, in terms of the mass transfer definition, $\langle\cinoutlet(t)\rangle$ plays the role of the 
characteristic concentration.
%This is the characteristic concentration in the definition, Eq. \ref{eq:main:definition}.  
The time-averaged
concentration flux ($\mathrm{\overline{Flux}}$) is calculated as the difference
between the overall average concentration in the whole domain
($\langle\coverall\rangle=\int_{V} C \mathrm{d}V /V$) at time $t_1$ and at time
$t_2$ divided by the time difference $t_2-t_1$. The agreement between numerical
simulations \cite{vanbaten-circular} and experimental correlations of \citet{bercic-mass}
% TODO: was good where?
was good. 
 \end{description}

The presented numerical approaches \cite{vanbaten-circular,kreutzer-overview}
can be criticized on a number of points. They mainly
relate to the bubble shape approximation, which is taken to be symmetrical, i.e.
consisting of two hemispheres and film for the case of flow in circular
capillaries. This is valid  for small capillary numbers only ($Ca<0.1$). As previously
discussed, for such capillary numbers the tracer is well mixed in the slug and 
the choice of the characteristic concentration needed for the mass transfer
coefficient, Eq. \ref{eq:main:definition}, is obvious. With minimal
differences in the results, it can either be the averaged concentration in the liquid slug
or the inlet/outlet space-averaged concentration. The latter is used in the formulation
of \citet{vanbaten-circular} presented above.

While it is clear that periodic boundary conditions can be employed for the calculation
 of hydrodynamic fields, the same does not apply to the mass transfer coefficient
simulations. Experimental correlations \cite{bercic-mass} show that the
concentration in a bubble train along the streamwise direction changes exponentially with distance.
Mass transfer simulations however, are made only for one unit cell
using periodic boundary conditions with the same concentration at the inlet
and at the outlet. The question is how single unit cell simulation corresponds to
experimental measurements arises where concentration difference is measured at
the distances of at least a few unit cells \cite{bercic-mass}. In
other words, one needs to understand how the discrete one unit cell simulation
corresponds to the continuous picture in experiments where one does not
distinguish discrete bubbles but takes measurements of concentration at different
locations.

Addressing situations for a rich number of hydrodynamic patterns, shapes, and
effects of bubble lengths, etc for bubble train flows, we feel that there is a need to examine 
carefully the strategies and assumptions behind the numerical calculations
of the mass transfer coefficient. We aim at establishing clear
procedures as to how properly obtain the mass transfer coefficient via a study
of different boundary conditions and different definitions of the characteristic concentration.
The case we want to examine is a two-dimensional bubble train flow between parallel plates.
We address the following issues:
 \begin{description}
 \item[I] Applicability of periodic boundary
conditions to determine the mass transfer coefficient when the vortex in the slug
disappears, i.e. when $Ca>0.7$.
 \item[II] Validity of the inlet/outlet-averaged or domain-averaged concentrations as
characteristic concentrations in the definition of the mass transfer
coefficient.
 \item[III] Translation of the continuous experimental picture to
numerical simulations of a few unit cells, the issue of correspondence
between space averages (simulations \cite{vanbaten-circular}) and time averages (experiment).
% \item[IV] Influence of the hydrodynamic pattern in flows with $Ca>0.7$ on the mass transfer.
\end{description}
In addition, at the end of the manuscript we present results of the dependence of the volumetric
mass transfer coefficient on the Peclet number that we compare with analytical \cite{irandoust} and experimental correlations
\cite{yue-mass}.  The thorough determination of the mass transfer coefficient and associated Sherwood number as a
function of other non-dimensional parameters such as gas holdup, bubble/slug
lenghts, and the capillary number is left for future studies.
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/benchmark_hydro.eps}
\caption{Simplified sketch of the bubble motion. Using periodic
conditions for the velocity field is natural, but needs evaluation for mass
transfer.
\label{fig:benchmark:hydro}}
\end{figure}

@@

Mass transfer problems in the lattice Boltzmann framework were mainly
addressed in a series of works of Ginzburg and co-authors
\cite{ginzburg-main,ginzburg-boundary-conditions,ginzburg-saturated-flow}.
In contrast to these works whose focus was on simulating the
advection-diffusion equation via the lattice Boltzmann framework, we concentrate
on the application side.  One should also mention the work of \citet{inamuro-scalar-boundary}
about heat and mass transfers in porous media and the work of
\citet{jos-mass} simulating lateral mixing in cross-channel flow.  The last two
works are focused on problems of homogeneous nature and do not provide
guidance as to how to obtain the mass transfer coefficient for heterogeneous cases.

The paper is organized as follows. We start with definitions of the volumetric mass transfer coefficient and apply them 
to the bubble train flow to derive expressions to connect the space- and time-averages.  Then, the lattice
Boltzmann model used to simulate mass transfer is presented, followed by benchmarks. Finally,
numerical simulations of various boundary conditions and simulations spanning a few unit cells for
different hydrodynamic patterns are presented to establish the procedure to determine the
volumetric mass transfer coefficient. The comparison with analytical correlations is also presented.

\section{Mass transfer definitions}
By definition, the mass transfer coefficient from a surface with an imposed constant
concentration $\cbubble$ is:
\beqal
\label{eq:main:definition}
&k_L=\frac{\dot{m}}{P \Delta C},\ \Delta C= \cbubble - \cmedium,
\feqal
where $\dot{m}$ is the mass flux $\Bigl[\frac{kg}{s}\Bigr]$, $P$ is the area of the surface
$\Bigl[m^2\Bigr]$, and $\Delta C$ is the concentration difference between the surface and the surrounding medium
$\Bigl[\frac{kg}{m^3}\Bigr]$. Therefore, $k_L$ has a dimension of 
$\Bigl[\frac{m}{s}\Bigr]$. Usually, the surrounding medium concentration is taken at an infinite distance
from the bubble. However, in the case of complicated geometries and non-homogeneous concentrations, 
the medium concentration can be the average concentration in the domain or the flux-averaged
concentration at the inlet or outlet, etc. Thus, one needs to establish a clear definition of  $\Delta C$ to determine the volumetric
mass transfer coefficient in the case of complex geometries and non-trivial hydrodynamic velocity patterns.

@@

%Another issue is that if the bubble is long enough then the film saturates with the tracer
%concentration and its influence on mass transfer can be negligible \cite{vanbaten-circular}.
Overall, mixing patterns of the film and liquid slugs are of great importance for the 
estimation of mass transfer \cite{yue-mass}. However, the assumptions usually taken for 
mass transfer calculations are small capillary numbers and certain mixing patterns such as to help to
estimate the mass transfer using the penetration theory of \citet{higbie}.

In comparison with analytical calculations and simplifications, the numerical approach can take into
account the complex mixing patterns and geometries. However, there are challenges as  how to
mimic the continuous picture where the medium is moving with bulk velocity $U=\ugas+\uliq$  as it is done in
experiments. Thus, the questions indicated in Section \ref{intro} arise.  The next section gives more
 details about numerical simulations.

\subsection{Numerical simulations}
\label{section:cases}
Ideally one wants to mimic the continuous picture as it is seen in experiments.
Thus, mass transfer simulations for a number of unit cells each containing a bubble are needed. As was
indicated above, there are two approaches towards it -- either to simulate the
bubble train and then to measure concentration along the pipe, Eq.~\ref{main:mass:transfer:expression},
or to transfer to the reference frame moving with the bulk
velocity $U$ and conduct the same measurements. However, both methods require tracking of
moving bubbles which is complicated from the numerical point of view. Therefore, one needs to come
up with a simple and smaller domain for calculations of the mass transfer coefficient, which closely mimics the
continuous picture of a large number of  separated bubbles.

@@

\section{Validation}
As was discussed earlier, analytical correlations for the mass transfer coefficient
have been derived for a Taylor bubble train flow  as two separate contributions: the mass transfer from two
half circles and the mass transfer from the film. We will examine these
mass transfer cases closely with the help of the lattice Boltzmann method and compare them
against analytical
solutions. The next sections will give a short introduction to the lattice Boltzmann method and present benchmark results.

@@

velocities $2,4,6,8,10,15,20,40$ times larger than the original velocities. The velocities and

@@

\ref{table:scaling:peclet} (top to bottom: 6 scalings, 5 scalings, 2 scalings, 1 scaling, 1 scaling). Some lines are indistinguishable showing that simulations with velocity scalings are consistent.  \label{fig:contours:scaling:peclet}}

@@

\section{Mass transfer for planar Poiseuille flow}
